---
title: 'DATA608: Assignment 3'
author: "Eric Lehmphul"
date: "10/7/2022"
output: 
  html_document:
    toc: true
    code_folding: show
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, warning=F, message=F}
library(shiny)
library(tidyverse)
library(plotly)
library(rsconnect)
```


```{r}
data <- read.csv("https://raw.githubusercontent.com/charleyferrari/CUNY_DATA_608/master/module3/data/cleaned-cdc-mortality-1999-2010-2.csv")
```

# Question 1

As a researcher, you frequently compare mortality rates from particular causes across different States. You need a visualization that will let you see (for 2010 only) the crude mortality rate, across all States, from one cause (for example, Neoplasms, which are effectively cancers). Create a visualization that allows you to rank States by crude mortality for each cause of death.


```{r}
# Get data from only 2010
data_2010 <- data %>% filter(Year == 2010)

# Create list of causes of death (ICD) to be used for a button
causes_of_death <- data_2010 %>% distinct(ICD.Chapter)
```




```{r}
# set up user interface
ui_q1 <- fluidPage(
  # Make Title
  titlePanel("Question 1: Mortality Rate By State"),
  
  # Set up cause of death selection button
   fluidRow(
     
      selectInput("ICD", 
                  label = "Select cause of death to be compared by State",
                  choices = causes_of_death$ICD.Chapter,
                  selected = "Neoplasms")
      
   ),
  plotlyOutput("bar_chart", height = 1000)
)
```


```{r}
server_q1 <- function(input, output){
  # Make data reactive
  data_2010_reactive <- reactive(data_2010 %>% filter(ICD.Chapter == input$ICD))
  
  # Generate graph
  output$bar_chart <- renderPlotly({plot_ly(data = data_2010_reactive(), x = ~Crude.Rate, y = ~State, type = "bar", orientation = "h", marker = list(color = "rgb(173, 177, 247)",
                            line = list(color = "rgb(20, 20, 20)",
                                        width = 1.5))) %>% plotly::layout(yaxis = list(categoryorder = "total ascending"))
    })
}
```


```{r}
shinyApp(ui = ui_q1, server = server_q1)
```




# Question 2

Often you are asked whether particular States are improving their mortality rates (per cause) faster than, or slower than, the national average. Create a visualization that lets your clients see this for themselves for one cause of death at the time. Keep in mind that the national average should be weighted by the national population.


Used the formula: $Crude.Rate = Deaths/Population * 100,000$ when grouped by `Year` and `ICD.Chapter` to calculate the national average. This formula was given in the help document under Crude Rate at the following link: https://wonder.cdc.gov/wonder/help/ucd.html#.

```{r}
# Prepare data for shiny app

# Create new variable, National.Avg
data_q2 <- data %>% group_by(Year, ICD.Chapter) %>%
  mutate(National.Avg = sum(Deaths) / sum(Population) * 100000)

```



```{r}
# Create user interface

ui_q2 <- fluidPage(
  titlePanel("Question 2: State Mortality Rate vs National Average"),
  
  fluidRow(
     
      selectInput("State", 
                  label = "Select State",
                  choices = unique(data_q2$State),
                  selected = "NY")
      
   ),
  
  fluidRow(
     
      selectInput("ICD", 
                  label = "Select cause of death",
                  choices = causes_of_death$ICD.Chapter,
                  selected = "Neoplasms")
      
   ),
  plotlyOutput("plot", height = 750)
)
```



```{r}
server_q2 <- function(input, output){
  data_q2_reactive <- reactive(data_q2)
  
  output$plot <- renderPlotly({
    data_q2_reactive() %>% filter(State == input$State, ICD.Chapter == input$ICD) %>% ungroup() %>%
    plot_ly(x = ~Year, y = ~Crude.Rate, type = 'bar', name = 'State Mortality Rate', marker = list(color = "rgb(60, 143, 57)",
                            line = list(color = "rgb(20, 20, 20)",
                                        width = 1.5))) %>%
  add_trace(x = ~Year, y = ~National.Avg, type = 'scatter',  mode = 'lines', name = 'National Mortality Rate', marker = list(color = "rgb(112, 23, 17)"), line = list(color = "rgb(112, 23, 17)", width = 2.5))

  })
}
```


```{r}
shinyApp(ui = ui_q2, server = server_q2)
```

